Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-10 paper/paper.tex
Fixed minor issues
Updated related work

\section{Related Work} 
\cred{Need to rewrite this entire section need to probably mention some of these 
papers at least i found by googling liveness based GC,\\
http://www.cs.hmc.edu/~oneill/papers/Simplifiers-MSPC.pdf\\
http://costa.ls.fi.upm.es/papers/costa/AlbertGGZ12.pdf\\
http://hirzels.com/martin/papers/ecoop01-gc-liveness.pdf\\
}
Previous attempts to  increase  the space
efficiency
of functional programs by additional  reclamation of memory fall in two
broad categories. In the first,  the program itself is instrumented to
manage  reclamation and reallocation  without the  aid of  the garbage
collector.    Such    attempts   include:   sharing    analysis   based
reallocation~\cite{jones89compile},                       deforestation
techniques~\cite{wadler88deforest,gill93ashort,chitil99deforest},
methods  based  on   linear  logic~\cite{hofmann00linear}  and  region
analysis~\cite{tofte98region}.   Closer  to  our approach,  there  are
methods   that  enable   the   garbage  collector   to  collect   more
garbage~\cite{inoue88analysis,lee05static}   by
explicitly  nullifying  pointers  that  are not  live.   However,  the
nullification,  done   at  compile  time,   requires  sharing  (alias)
analysis.  Our  method, in contrast, does not  require sharing because
of the availability of the heap  itself at runtime.  To the best of our
knowledge, this is the first  attempt at liveness-based marking of the
heap during garbage collection.

@@

  Orthogonally,  we plan  to improve  the efficiency  of the  liveness
  based garbage collector using heuristics  such as limiting the depth
  of  DFAs,   merging  nearly-equivalent   states  and   using  better
  representation  and algorithms  for automata  manipulation. We  also
  need  to   investigate  the  interaction  of   liveness  with  other
  collection schemes, such as incremental and generational collection.
  In short,  we need to  investigate the  ways to make  liveness based
  garbage collection attractive for practical collectors.  }

@@

\subsection{References}
\bibliography{fun_hra}{}
\bibliographystyle{abbrv}

>>>>

\section{Related Work} 
Although augmenting  garbage collection with liveness  information has
been  studied earlier,  it was  mainly  in the  context of  imperative
languages~\cite{Albert,  Hirzel}.  Due  to   the  presence  of  global
variables  and mutation,  practical utility  of simple  liveness based
techniques are found to be inefficient.

In the space  of functional languages, improving  space efficiency has
been  mainly  studied  as  a compile  time  activity.  Either  through
rewriting                methods                such                as
deforestation~\cite{wadler88deforest,gill93ashort,chitil99deforest},
sharing  analysis   based  reallocation~\cite{jones89compile},  region
based  analysis~\cite{tofte98region} or  through insertion  of compile
time              nullifying              statements              such
as~\cite{inoue88analysis,lee05static,Hamilton}.    The  compile   time
marking approaches all rely on an efficient and precise alias analysis
and in the absence of it cannot provide notable improvement.

The closest  to our approach  is the liveness based  garbage collector
implemented in~\cite{asati14lgc}. We extend  their work to handle lazy
evaluation and closures.

%% A similar approach has been suggested in~\cite{ONeill} where they suggest augmenting garbage collection with extra information using {\it simplifiers}.

%% Previous  attempts  to increase  the  space  efficiency of  functional
%% programs  by  additional  reclamation  of memory  fall  in  two  broad
%% categories. In the first, the program itself is instrumented to manage
%% reclamation and reallocation without the aid of the garbage collector.
%% Such      attempts      include:      sharing      analysis      based
%% reallocation~\cite{jones89compile},                      deforestation
%% techniques~\cite{wadler88deforest,gill93ashort,chitil99deforest},
%% methods  based  on   linear  logic~\cite{hofmann00linear}  and  region
%% analysis~\cite{tofte98region}.   Closer  to  our approach,  there  are
%% methods   that  enable   the   garbage  collector   to  collect   more
%% garbage~\cite{inoue88analysis,lee05static}  by  explicitly  nullifying
%% pointers  that are  not  live.  However,  the  nullification, done  at
%% compile  time,  requires sharing  (alias)  analysis.   Our method,  in
%% contrast, does not require sharing  because of the availability of the
%% heap itself  at runtime.  To  the best of  our knowledge, this  is the
%% first attempt  at liveness-based  marking of  the heap  during garbage
%% collection.

@@

  Orthogonally,  we plan  to improve  the efficiency  of the  liveness
  based garbage collector using heuristics  such as limiting the depth
  of  DFAs,   merging  nearly-equivalent   states  and   using  better
  representation  and algorithms  for automata  manipulation. We  also
  need  to   investigate  the  interaction  of   liveness  with  other
  collection schemes, such as incremental and generational collection.
  In  summary, we  need to  investigate  ways to  make liveness  based
  garbage collection attractive for practical collectors.  }

@@

\bibliography{fun_hra}{}
\bibliographystyle{abbrv}
